On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-08-29 paper.tex
Explained more philosophy

\section{A variant of Rehmann theorem}
Let $A$ be arbitrary commutative unital ring and $\mathfrak{m}$ be its ideal.
Denote by $B$ the ring $\mathfrak{m}[t^{-1}] + A[t]$ with the obvious $\mathbb{Z}$-grading.
In other words, as a module $B$ decomposes as a direct sum of its homogeneous components $\oplus_k B_k$,
 where each $B_k$ is either $A\cdot t^k$ for $k \geq 0$ or $\mathfrak{m} \cdot t^k$ for $k<0$.
%If $a$ is a homogeneous element of $B$ (i.\,e. an element of some $B_k$) we denote its homogeneous degree by $|a|$.

Let $\Phi$ be a reduced irreducible root system of rank $\geq 3$.
The aim of this section is to find a presentation of the Steinberg group $\St(\Phi, B)$ in terms of generators $x_{\alpha}(\xi)$
 for which $\xi$ is homogeneous of degree $\leq 1$, see~\cref{prop:tul3.3} below.
Our presentation should be thought of as an analogue of~\cite[Lemma~3.3]{Tu83} in the case $n=4$ (with variables $t$ and $t^{-1}$ swapped).
Notice that Tulenbaev's short proof of \cite[Lemma~3.3]{Tu83} which works for $\Phi=\rA_\ell$, $\ell\geq 4$ does not generalize to the case $\ell=3$,
%TODO: Explain why we are dealing with general \Phi rather than just the case \rA_3 -- Reason the difficulty is the same
 so we are forced to recourse to a much longer argument inspired by the proof of Rehmann--Soul{\'e} theorem on
 finite presentation of Steinberg groups over polynomial rings (cf.~\cite{Re75, RS76}).

@@

\begin{comment} 
\begin{rem}
There are several reasons why we can not simply refer to~\cite{Re75} or~\cite{RS76} and need to reproduce arguments from these papers.
\begin{itemize}
 \item Rehmann and Soul{\'e} deal with usual polynomial rings rather than $\mathbb{Z}$-graded rings
  (i.\,e. their results correspond to the special case $\mathfrak{m}=0$ of our result);
 \item Presentations from~\cite{Re75,RS76} are formulated in terms of generating set $\XX{2}$ which is not sufficient for our purposes
  (we need a presentation in terms of $\XX{1}$);
 \item It is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}, while we work with
  arbitrary coefficient rings $A$.
\end{itemize} 
\end{rem}
\end{comment} %TODO: Find proper place for this discussion - it may not be appropriate here

For $m\geq 2$ define $G_m$ as the group presented by generators $\mathcal{X}_m$ and the following three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_{\alpha}(\xi) x_{\alpha}(\eta) & = x_{\alpha}(\xi+\eta),&  \xi,\eta\in B_d,\ d\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\alpha'}(\eta))] &  = 1, & \text{in the case $\alpha+\alpha'\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_d,\ \eta \in B_e,\ d,e\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\alpha'}(\eta)] & = x_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi\eta), & \text{in the case $\alpha+\alpha'\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_d,\ \eta \in B_e,\ d, e, d+e \leq m.
 \end{align}
For $m=1$ define $G_1$ as the group presented by generators $\mathcal{X}_1$, 
three above families of relations $(a_1), (b_1), (c_1)$ and the following additional family of relations:
\begin{align}  \label{eq:d1} \tag{$d_1$} [[x_{\alpha+\beta+\gamma}(\xi), x_{-\gamma}(\eta)], x_{\beta+\gamma}(\zeta) ] & = 1, 
 & \xi,\eta,\zeta\in B_1,\ \text{ $(\alpha, \beta, \gamma)$ is an $\rA_3$-triple. }
\end{align}
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

The obvious embedding of generators $\mathcal{X}_m \subseteq \mathcal{X}_{m+1}$ induces a map $f_m\colon G_m \to G_{m+1}$.
Our goal is to show the following result.
\begin{prop}\label{prop:tul3.3}
 For $m\geq 1$ the map $f_m$ is an isomorphism. Consequently, $\St(\Phi, B)$ can be presented by generators $\mathcal{X}_1$
  and relations $(a_1)$, $(b_1)$, $(c_1)$, $(d_1)$.
\end{prop}
%TODO: Deduce 2nd assertion from 1st
Notice that the second assertion follows from the first one.
Indeed, this would give the required isomorphism $G_1 \cong \St(\Phi, B)$ since $\langle \XX{\infty} \mid \RR{\infty} \rangle$ is evidently isomorphic to $\St(\Phi, B)$.
%Denote by $F(S)$ the free group on the generating set $S$.
%It is clear that $\RR{n}\subseteq\RR{n+1}$, moreover $\langle \RR{n} \rangle = \langle \RR{\infty}\rangle \cap F(\XX{n})$.
%well, not so clear, actually :)

@@

>>>>

\section{A variant of Rehmann theorem}
Let $A$ be arbitrary commutative unital ring and $\mathfrak{m}$ be its ideal.
Denote by $B$ the ring $\mathfrak{m}[t^{-1}] + A[t]$ with the obvious $\mathbb{Z}$-grading.
In other words, as a module $B$ decomposes into a direct sum of its homogeneous components $\oplus_k B_k$,
 where each $B_k$ is either $A\cdot t^k$ for $k \geq 0$ or $\mathfrak{m} \cdot t^k$ for $k<0$.

%If $a$ is a homogeneous element of $B$ (i.\,e. an element of some $B_k$) we denote its homogeneous degree by $|a|$.

Let $\Phi$ be a reduced irreducible root system of rank $\geq 3$.
The main result of this section is~\cref{prop:tul3.3} which gives a presentation of Steinberg group $\St(\Phi, B)$ in terms of generating set $\mathcal{X}_1$.
In the sequel we will need this presentation only in the special case $\Phi=\rA_3$. However, we preferred to prove a more general statement since the proof in the special case $\Phi=\rA_3$ has almost the same length.

This presentation give by~\cref{prop:tul3.3} can be thought of as an analogue of~\cite[Lemma~3.3]{Tu83} in the case $n=4$ (with variables $t$ and $t^{-1}$ swapped).
Notice that Tulenbaev's short proof of \cite[Lemma~3.3]{Tu83} in the case $\Phi=\rA_\ell$, $\ell\geq 4$ does not work in the case $\Phi=\rA_3$,
%TODO: Explain why we are dealing with general \Phi rather than just the case \rA_3 -- Reason the difficulty is the same
 which forces us to recourse to a much longer argument inspired by the proof of Rehmann--Soul{\'e} theorem on
 finite presentation of Steinberg groups over polynomial rings (cf.~\cite{Re75, RS76}).

\begin{rem}
There are several reasons why we can not simply refer to~\cite{Re75} or~\cite{RS76} and instead need to reprove everything by hand.
\begin{itemize}
 \item Rehmann and Soul{\'e} deal with usual polynomial rings rather than $\mathbb{Z}$-graded rings
  (i.\,e. their results correspond to the special case $\mathfrak{m}=0$ of our result);
 \item Presentations from~\cite{Re75,RS76} are given in terms of generating set $\XX{2}$ rather than $\mathcal{X}_1$ which is not precise enough for our purposes;
 \item It is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}, while we must work with arbitrary coefficient ring $A$.
\end{itemize} 
\end{rem}

@@

For $m\geq 2$ define $G_m$ as the group presented by generators $\mathcal{X}_m$ and the following three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_{\alpha}(\xi) x_{\alpha}(\eta) & = x_{\alpha}(\xi+\eta),&  \xi,\eta\in B_d,\ d\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\alpha'}(\eta))] &  = 1, & \text{in the case $\alpha+\alpha'\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_d,\ \eta \in B_e,\ d,e\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\alpha'}(\eta)] & = x_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi\eta), & \text{in the case $\alpha+\alpha'\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_d,\ \eta \in B_e,\ d, e, d+e \leq m.
 \end{align}
For $m=1$ define $G_1$ as the group presented by generators $\mathcal{X}_1$, 
three above families of relations $(a_1), (b_1), (c_1)$ and the following additional family:
\begin{align}  \label{eq:d1} \tag{$d_1$} [[x_{\alpha+\beta+\gamma}(\xi), x_{-\gamma}(\eta)], x_{\beta+\gamma}(\zeta) ] & = 1, 
 & \xi,\eta,\zeta\in B_1,\ \text{ $(\alpha, \beta, \gamma)$ is an $\rA_3$-triple. }
\end{align}
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

The obvious embedding of generators $\mathcal{X}_m \subseteq \mathcal{X}_{m+1}$ induces a map $f_m\colon G_m \to G_{m+1}$.
Our goal is to show the following result.
\begin{prop}\label{prop:tul3.3}
 For $m\geq 1$ the map $f_m$ is an isomorphism. Consequently, $\St(\Phi, B)$ can be presented by generators $\mathcal{X}_1$
  and relations $(a_1)$, $(b_1)$, $(c_1)$, $(d_1)$.
\end{prop}
\begin{rem}
 The above presentation seems to be the shortest possible in the cases $\Phi=\rA_3, \rD_\ell$. On the other hand, the analogue of the above proposition for $\Phi=\rA_\ell$, $\ell\geq 4$ and $\Phi=\rE_\ell$ holds without additional relation $(d_1)$. For $\rA_\ell$ this is proved in~\cite[Lemma~3.3]{Tu83}, for $\rE_\ell$ this can be easily deduced using the amalgamation technique (see~???).
\end{rem}

@@

\begin{proof}[Proof of~\cref{prop:tul3.3}]
 %TODO: Deduce 2nd assertion from 1st
Notice that the second assertion follows from the first one.
Indeed, this would give the required isomorphism $G_1 \cong \St(\Phi, B)$ since $\langle \XX{\infty} \mid \RR{\infty} \rangle$ is evidently isomorphic to $\St(\Phi, B)$.
%Denote by $F(S)$ the free group on the generating set $S$.
%It is clear that $\RR{n}\subseteq\RR{n+1}$, moreover $\langle \RR{n} \rangle = \langle \RR{\infty}\rangle \cap F(\XX{n})$.
%well, not so clear, actually :)

\end{proof}
